Hydra Battles and Cie

### 2021-05-17 doc/part-hydras.tex
Corrected references to the Hardy hierarchy
The variant of the hardy hierarchi is now called H'_ alpha, in order to avoid
confusion with the litterature

\begin{remark}
Some notations (mainly names of fast-growing functions) of our development may differ slightly from the litterature. Although this fact does not affect our proofs, we are preparing a future 
version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles. This change will be provisionally made in a branch called \texttt{Hardy} (in reference to the Hardy hierarchy).
\end{remark}
%\section{Introduction}

@@

\section{The Wainer-Hardy hierarchy (functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}

@@

\begin{remark}
  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions. In the future versions of this development, we will correct the references to the litterature.
\end{remark}

@@

For each ordinal $\alpha$ below $\epsilon_0$, $H_\alpha$ is 
a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:

@@

\begin{itemize}
\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
\item If $\alpha=\textrm{succ}(\beta)$, then 
$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
\item If $\alpha$ is a limit ordinal, then 
$H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
\end{itemize}

\subsection{Hardy functions in \texttt{Coq}}

We define a function \texttt{H\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.H_alpha.html\#H_}{Epsilon0.H\_alpha}}

\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy)}
\index{coq}{Plug-ins!Equations}
\label{Functions:H-alpha}

\begin{Coqsrc}
Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
  H_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  H_ (Canon alpha (S i))  i ;
            | right notlimit =>  H_ (Pred alpha) (S i)}}.

@@

\begin{Coqsrc}
Lemma H_eq1 : forall i, H_ Zero i = i.
Proof.   intro i; now rewrite H__equation_1.  Qed.

Lemma H_eq2 alpha i : Is_Succ alpha ->
                      H_ alpha i = H_ (Pred alpha) (S i).

Lemma H_eq3 alpha i : Limitb alpha ->
                      H_ alpha i =  H_ (Canon alpha (S i)) i.

Lemma H_eq4  alpha i :  H_ (Succ alpha) i = H_ alpha (S i).
\end{Coqsrc}

\subsection{First  steps of the Hardy hierarchy}
Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.

@@

By induction on $i$, we prove a simple expression of \texttt{H\_ (Fin i)}, where 
\texttt{Fin $i$}  is the $i$-th finite ordinal.

\begin{Coqsrc}
Lemma H_Fin : forall i k: nat,  H_ (Fin i) k = (i+k)%nat.
Proof with eauto with E0.
  induction i.
  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
  - intros ;simpl; autorewrite with H_rw E0_rw ... 
    rewrite IHi; lia. 
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
Proof with auto with E0.
  intro k; rewrite H_eq3 ...
  - replace (Canon omega (S k)) with (Fin (S k)).
    + rewrite H_Fin; lia.
    +  now autorewrite with E0_rw.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H_Plus_Fin alpha : forall i k : nat,
    H_ (alpha + i)%e0 k = H_ alpha (i + k)%nat.
(* Proof by induction on i *)
\end{Coqsrc}

Then, we get easily formulas for $H_{\omega+i}$, and $H_{\omega\times i}$ for any natural number $i$.

\begin{Coqsrc}
Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
Proof.
 rewrite H_lim_eqn; simpl Canon.
 - ochange  (CanonS  (omega * 2)%e0 k)  (omega + (S k))%e0.
  + rewrite H_Plus_Fin, H_omega;  lia.
  -  now compute.
Qed.

Lemma H_omega_3 k : H_ (omega * 3)%e0 k = (8 * k + 7)%nat.

Lemma H_omega_4 k : H_ (omega * 4)%e0 k = (16 * k + 15)%nat.

Lemma H_omega_i i  : forall k,
    H_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
\end{Coqsrc}

Crossing a new limit, we prove the following equality: 
$$H_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.

Lemma H_omega_sqr : forall k,
    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
Proof.
  intro k; 
   rewrite H_lim_eqn; auto with E0.
  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
    +  rewrite H_omega_i; simpl (exp2 (S k)).
       *  rewrite Nat.add_pred_r.
          -- lia. 
          --   generalize (exp2_not_zero k);  lia.
    + cbn; f_equal; lia.
Qed.
\end{Coqsrc}

@@

Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
Since the canonical sequence of $\omega^\omega$ is composed of all the
$\omega^i$, we first need to express $H_{\omega^i}$ for any natural number $i$.

@@

\begin{enumerate}
\item First,  we express $H_{\omega^\alpha \times (i+2)}$ in terms of
$H_{\omega^\alpha \times (i+1)}$.

\begin{Coqsrc}
Lemma H_Omega_term_1 : alpha <> Zero -> forall  k,  
    H_ (Omega_term alpha (S i)) k =
    H_ (Omega_term alpha i) (H_ (Phi0 alpha) k).
\end{Coqsrc}

\item
Then, we prove by induction on $i$ that $H_{\omega^\alpha \times (i+1)}$ is just the
$(i+1)$-th iterate of $H_{\omega^\alpha}$.

\begin{Coqsrc}
Lemma H_Omega_term (alpha : E0)  :
alpha <> Zero -> forall i k, 
  H_ (Omega_term alpha i) k = iterate  (H_ (Phi0 alpha)) (S i) k.
\end{Coqsrc}

\item In particular, we have got a formula for $H_{\omega^{i+1}}$.

\begin{Coqsrc}
Definition H_succ_fun f k := iterate f (S k) k.

Lemma H_Phi0_succ alpha  : alpha <> Zero -> forall k,
      H_ (Phi0 (Succ alpha)) k = H_succ_fun (H_ (Phi0 alpha)) k.

Lemma H_Phi0_Si : forall i k,
      H_ (Phi0 (S i)) k = iterate H_succ_fun i (H_ omega) k. 
\end{Coqsrc}

\end{enumerate}
We get now a  formula for $H_{\omega^3}$:

\begin{Coqsrc}
Lemma H_omega_cube : forall k,
    H_ (Phi0 3)%e0 k = iterate (H_ (Phi0 2))  (S k) k.
Proof.
  intro k; rewrite <-FinS_eq, -> Fin_Succ, H_Phi0_succ; auto.
  compute; injection 1; discriminate.
Qed.
\end{Coqsrc}

@@

It seems hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example, the number $H_{\omega^3}(3)$.

\begin{Coqsrc}
Section H_omega_cube_3.

@@

Remark R0 k :  H_ (Phi0 3)%e0 k = iterate f (S k) k.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fact F0 : H_ (Phi0 3) 3 = f (f (f (f 3))).
 rewrite R0; reflexivity. 
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fact F1 : H_ (Phi0 3) 3 = f (f N).
Proof.
 rewrite H_omega_cube_0; reflexivity. 
Qed.

Lemma F1_simpl : H_ (Phi0 3) 3 =
                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.

@@

{\Large
$$
H_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
$$
}

We leave as an exercise to determine the best approximation as possible of
 the size of this number (for instance its number of digits).  For instance, if
we do not take into account the multiplications in the formula above,
we obtain that, in base $2$, the number $H_{\omega^3}(3)$ has at least
$2^{10^{21}}$  digits. But it is still an under-approximation !

\begin{Coqsrc}
End H_omega_cube_3.
\end{Coqsrc}

Now, we have got at last an exact formula for $H_{\omega^\omega}$.

\begin{Coqsrc}
Lemma H_Phi0_omega : forall k, H_ (Phi0 omega) k =
                               iterate H_succ_fun  k (H_ omega) k.
Proof with auto with E0.
  intro k; rewrite H_lim_eqn, <- H_Phi0_Si ...
  -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H_Phi0_omega_closed_formula k :
  H_ (Phi0 omega) k =
  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
               k
               (fun k : nat => S (2 * k)%nat)
               k.
\end{Coqsrc}

@@

\subsection{Abstract properties of Hardy functions}
~\label{sect:H-alpha-prop}

Since pure computation seems to be useless for dealing with expressions of the form $H_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H_\alpha(k)$ and $H_\beta(l)$, in terms of comparison
between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.

But beware of non-theorems! For instance, one could believe that $H$ is monotonous in its first argument. The following proof shows this is false.

\begin{Coqsrc}
Remark H_non_mono1 :
  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
                          (H_ alpha k <= H_ beta k)%nat).
Proof.
 intros H ;specialize (H 42 omega 3).
 assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
 apply H in H0; rewrite H_Fin, H_omega  in H0; lia.
Qed.
\end{Coqsrc}

On the contrary, the fonctions of the Hardy hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
\begin{itemize}
\item the function $H_\alpha$ is strictly monotonous :
      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$

\item For any $n\geq 1$, $H_\alpha(n)<H_{\alpha+1}(n)$.
\emph{We say that $H_{\alpha+1}$ dominates $H_\alpha$ from $1$}.
\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
$H_\beta(n)\leq H_\alpha(n)$.
\end{itemize}

@@

\begin{Coqsrc}
Section Proof_of_Abstract_Properties.
  Record P (alpha:E0) : Prop :=
    mkP {
        PA : strict_mono (H_ alpha);
        PB : alpha <> Zero -> forall n,  (n < H_ alpha n)%nat;
        PC : H_ alpha <<= H_ (Succ alpha);
        PD : dominates_from 1 (H_ (Succ alpha)) (H_ alpha);
        PE : forall beta n, Canon_plus n alpha beta -> 
                            (H_ beta n <= H_ alpha n)%nat}.

@@

\subsection{Comparison between \texttt{L\_} and \texttt{H\_} }

By well-founded induction on $\alpha$, we prove that our $L$ hierachy is ``almost'' the Hardy hierarchy (up to a small shift).

\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H_L_}{Epsilon0.L\_alpha}}

\begin{Coqsrc}
 Theorem H_L_ alpha: forall i:nat,  (H_ alpha i <= L_ alpha (S i))%nat.
\end{Coqsrc}

@@

The following theorem relates the length of (standard) battles with the Hardy hierarchy.

@@

\begin{Coqsrc}
Theorem battle_length_std_Hardy (alpha : E0) :
  alpha <> Zero ->
  forall k , 1 <= k -> exists l: nat,  
       H_ alpha k - k <= l /\
       battle_length standard k (iota (cnf alpha)) l.    
\end{Coqsrc}

@@

\index{hydras}{Exercises}
\begin{exercise}
Let us quote a theorem from ~\cite{KS81} (page 297).

\begin{quote}
\begin{align*}
  H_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
 F_{\alpha}(n+1) &\geq H_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
\end{align*}
Thus $H_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.

\end{quote}

 But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper ! Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.

>>>>

\begin{remark}
Some notations (mainly names of fast-growing functions) of our development may differ slightly from the litterature. Although this fact does not affect our proofs, we are preparing a future 
version where the names $F\_alpha$, $f\_alpha$, $H\_alpha$, etc., are fully consistent with the cited articles. 
\end{remark}
%\section{Introduction}

@@

\section{A variant of the Wainer-Hardy hierarchy}

@@

@@

\begin{remark}
  Indeed, the functions presented in this section are a \emph{variant} of the Hardy hierarchy of functions. In the future versions of this development, we will correct the references to the litterature. For the time being, we call our functions $H'\_alpha$ in order to underline the difference from ``classic'' Hardy functions.
\end{remark}

For each ordinal $\alpha$ below $\epsilon_0$, $H'_\alpha$ is 
a total arithmetic function, defined  by  transfinite recursion on $\alpha$, according to three cases:

@@

\begin{itemize}
\item If $\alpha=0$, then $H'_\alpha (k)= k$ for any natural number $k$.
\item If $\alpha=\textrm{succ}(\beta)$, then 
$H'_\alpha(k)=H'_\beta(k+1)$ for any $k \in \mathbb{N}$
\item If $\alpha$ is a limit ordinal, then 
$H'_\alpha(k) = H'_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
\end{itemize}

\begin{remark}
 The ``classic'' definition of the Wainer-Hardy hierarchy differs in the third equation.

\begin{itemize}
\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
\item If $\alpha=\textrm{succ}(\beta)$, then 
$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
\item If $\alpha$ is a limit ordinal, then 
$H_\alpha(k) = H_{(\canonseq{\alpha}{k})}(k)$ for any $k\in \mathbb{N}$.
\end{itemize}

\subsection{Definition in \texttt{Coq}}

We define a function \texttt{H'\_} of type \texttt{E0 -> nat -> nat} by transfinite induction over the type \texttt{E0} of the well formed ordinals below $\epsilon_0$.

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hprime.html\#H_}{Epsilon0.Hprime}}

\index{hydras}{Library Epsilon0!Functions!H\_@H\_ (Hardy hierarchy (variant))}
\index{coq}{Plug-ins!Equations}
\label{Functions:Hprime-alpha}

\begin{Coqsrc}
Equations H'_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  H'_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  H'_ (Canon alpha (S i))  i ;
            | right notlimit =>  H'_ (Pred alpha) (S i)}}.

@@

\begin{Coqsrc}
Lemma H'_eq1 : forall i, H'_ Zero i = i.
Proof.
  intro i; now rewrite H'__equation_1. 
Qed.

Lemma H'_eq2 alpha i : Succb alpha ->
                       H'_ alpha i = H'_ (Pred alpha) (S i).

Lemma H'_eq3 alpha i : Limitb alpha ->
                      H'_ alpha i =  H'_ (Canon alpha (S i)) i.

Lemma H'_succ_eqn  alpha i :  
            H'_ (Succ alpha) i = H'_ alpha (S i).
\end{Coqsrc}

\subsection{First  steps of the H' hierarchy}
Using rewrite rules from \texttt{H'\_eq1} to \texttt{H'\_succ\_eqn}, we can explore the functions $H'_\alpha$ for  small values of $\alpha$.

%ICI

@@

By induction on $i$, we prove a simple expression of \texttt{H'\_ (Fin i)}, where 
\texttt{Fin $i$}  is the $i$-th finite ordinal.

\begin{Coqsrc}
Lemma H'_Fin : forall i k : nat,  H'_  (Fin i) k = (i+k)%nat.
Proof with eauto with E0.
  induction i.
  - intros; simpl Fin; simpl; autorewrite with H'_rw E0_rw ... 
  - intros ;simpl; autorewrite with H'_rw E0_rw ... 
    rewrite IHi; abstract lia. 
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_omega : forall k, H'_ omega k = S (2 * k)%nat.
Proof with auto with E0.
  intro k; rewrite H'_eq3 ...
  - replace (Canon omega (S k)) with (Fin (S k)).
    + rewrite H'_Fin; abstract lia.
    + now autorewrite with E0_rw.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_Plus_Fin alpha : forall i k : nat,
    H'_ (alpha + i)%e0 k = H'_ alpha (i + k)%nat.
(* Proof by induction on i *)
\end{Coqsrc}

Then, we get easily formulas for $H'_{\omega+i}$, and $H'_{\omega\times i}$ for any natural number $i$.

\begin{Coqsrc}
Lemma H'_omega_double k : H'_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
Proof.
  rewrite H'_eq3; simpl Canon.
 -   ochange  (CanonS  (omega * FinS 1)%e0 k)  (omega + (S k))%e0.
  + rewrite H'_Plus_Fin, H'_omega;  abstract lia.
  -  now compute.
Qed.

Lemma H'_omega_3 k : H'_ (omega * 3)%e0 k = (8 * k + 7)%nat.

Lemma H'_omega_4 k : H'_ (omega * 4)%e0 k = (16 * k + 15)%nat.

Lemma H'_omega_i i  : forall k,
    H'_ (omega * i)%e0 k = (exp2 i * k + Nat.pred (exp2 i))%nat.
\end{Coqsrc}

Crossing a new limit, we prove the following equality: 
$$H'_{\omega^2} (k) = 2 ^ {k+1} \times (k+1) - 1$$.

\begin{Coqsrc}
Lemma H'_omega_sqr : forall k,
    H'_ (Phi0  2)%e0 k = (exp2 (S k ) * (S k) - 1)%nat.
Proof.
  intro k; rewrite H'_eq3; auto with E0.
  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
    +  rewrite H'_omega_i; simpl (exp2 (S k)).
       *  rewrite Nat.add_pred_r.
          -- abstract lia. 
          --   generalize (exp2_not_zero k);  abstract lia.
    + cbn; f_equal; abstract lia.
Qed.
\end{Coqsrc}

@@

Our next step would be to prove an exact formula for $H'_{\omega^\omega}(k)$.
Since the canonical sequence of $\omega^\omega$ is composed of all the
$\omega^i$, we first need to express $H'_{\omega^i}$ for any natural number $i$.

@@

\begin{enumerate}
\item First,  we express $H'_{\omega^\alpha \times (i+2)}$ in terms of
$H'_{\omega^\alpha \times (i+1)}$.

\begin{Coqsrc}
Lemma H'_Omega_term_1 : alpha <> Zero -> forall  k,  
    H'_ (Omega_term alpha (S i)) k =
    H'_ (Omega_term alpha i) (H'_ (Phi0 alpha) k).
\end{Coqsrc}

\item
Then, we prove by induction on $i$ that $H'_{\omega^\alpha \times (i+1)}$ is just the
$(i+1)$-th iterate of $H'_{\omega^\alpha}$.

\begin{Coqsrc}
Lemma H'_Omega_term (alpha : E0)  :
alpha <> Zero -> forall i k, 
  H'_ (Omega_term alpha i) k = iterate  (H'_ (Phi0 alpha)) (S i) k.
\end{Coqsrc}

\item In particular, we have got a formula for $H'_{\omega^{i+1}}$.

\begin{Coqsrc}
Definition H'_succ_fun f k := iterate f (S k) k.

Lemma H'_Phi0_succ alpha  : alpha <> Zero -> forall k,
      H'_ (Phi0 (Succ alpha)) k = H'_succ_fun (H'_ (Phi0 alpha)) k.

Lemma H'_Phi0_Si : forall i k,
      H'_ (Phi0 (S i)) k = iterate H'_succ_fun i (H'_ omega) k. 
\end{Coqsrc}

\end{enumerate}
We get now a  formula for $H'_{\omega^3}$:

\begin{Coqsrc}
Lemma H'_omega_cube : forall k,
    H'_ (Phi0 3)%e0 k = iterate (H'_ (Phi0 2))  (S k) k.
Proof.
  intro k; rewrite <-FinS_eq, -> Fin_Succ, H'_Phi0_succ; auto.
  compute; injection 1; discriminate.
Qed.
\end{Coqsrc}

@@

It seems hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example: the number $H'_{\omega^3}(3)$.

\begin{Coqsrc}
Section H'_omega_cube_3.

@@

Remark R0 k :  H'_ (Phi0 3)%e0 k = iterate f (S k) k.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fact F0 : H'_ (Phi0 3) 3 = f (f (f (f 3))).
 rewrite R0; reflexivity. 
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fact F1 : H'_ (Phi0 3) 3 = f (f N).
Proof.
 rewrite F0; reflexivity. 
Qed.

Lemma F1_simpl : H'_ (Phi0 3) 3 =
                 (exp2 (exp2 (S N) * S N) * (exp2 (S N) * S N) - 1)%nat.

@@

{\Large
$$
H'_{\omega^3}(3) =  2 ^ {(2 ^ {N + 1} \, (N+1) )}   \,  (2 ^ {N+1} \, ( N +1) ) - 1
$$
}

We leave as an exercise to determine the best approximation as possible of
 the size of this number (for instance its number of digits).  For instance, if
we do not take into account the multiplications in the formula above,
we obtain that, in base $2$, the number $H'_{\omega^3}(3)$ has at least
$2^{10^{21}}$  digits. But it is still an under-approximation !

\begin{Coqsrc}
End H'_omega_cube_3.
\end{Coqsrc}

Now, we have got at last an exact formula for $H'_{\omega^\omega}$.

\begin{Coqsrc}
Lemma H'_Phi0_omega : forall k, H'_ (Phi0 omega) k =
                               iterate H'_succ_fun  k (H'_ omega) k.
Proof with auto with E0.
  intro k; rewrite H'_eq3, <- H'_Phi0_Si ...
  -  rewrite CanonS_Canon, CanonS_Phi0_lim;  f_equal ...
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H'_Phi0_omega_closed_formula k :
  H'_ (Phi0 omega) k =
  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
               k
               (fun k : nat => S (2 * k)%nat)
               k.
\end{Coqsrc}

@@

\subsection{Abstract properties of  
\texorpdfstring{$H'_\alpha$}{H'}}
~\label{sect:H-alpha-prop}

Since pure computation seems to be useless for dealing with expressions of the form $H'_\alpha(k)$, even for small values of $\alpha$ and $k$, we need to prove theorems for comparing $H'_\alpha(k)$ and $H'_\beta(l)$, in terms of comparison
between $\alpha$ and $\beta$ on the one hand, $k$ and $l$ on the other hand.

But beware of non-theorems! For instance, one could believe that $H'$ is monotonous in its first argument. The following proof shows this is false.

\begin{Coqsrc}
Remark H'_non_mono1 :
  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
                          (H'_ alpha k <= H'_ beta k)%nat).
Proof.
 intros H ;specialize (H 42 omega 3).
 assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
 apply H in H0; rewrite H'_Fin, H'_omega  in H0; lia.
Qed.
\end{Coqsrc}

On the contrary, the fonctions of the $H'$ hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
\begin{itemize}
\item the function $H'_\alpha$ is strictly monotonous :
      For all $n,p \in\mathbb{N}, n < p \Rightarrow H'_\alpha(n)< H'_\alpha(p)$.
\item If $\alpha \not= 0$, then for every $n$, $n<H'_\alpha(n)$.
\item The function $H'_\alpha$ is pointwise less or equal than $H'_{\alpha+1}$

\item For any $n\geq 1$, $H'_\alpha(n)<H'_{\alpha+1}(n)$.
\emph{We say that $H'_{\alpha+1}$ dominates $H'_\alpha$ from $1$}.
\item For any $n$ and $\beta$, if $\alpha \xrightarrow[n]{} \beta$, then
$H'_\beta(n)\leq H'_\alpha(n)$.
\end{itemize}

@@

\begin{Coqsrc}
Section Proof_of_Abstract_Properties.
  Record P (alpha:E0) : Prop :=
    mkP {
        PA : strict_mono (H'_ alpha);
        PB : alpha <> Zero -> forall n,  (n < H'_ alpha n)%nat;
        PC : H'_ alpha <<= H'_ (Succ alpha);
        PD : dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha);
        PE : forall beta n, Canon_plus n alpha beta -> 
                            (H'_ beta n <= H'_ alpha n)%nat}.

@@

\subsection{Comparison between \texttt{L\_} and \texttt{H'\_} }

By well-founded induction on $\alpha$, we prove a simple relation between $L_\alpha$ and $H'\_alpha$.

\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html\#H'_L_}{Epsilon0.L\_alpha}}

\begin{Coqsrc}
 Theorem H'_L_ alpha: forall i:nat,  (H'_ alpha i <= L_ alpha (S i))%nat.
\end{Coqsrc}

@@

The following theorem relates the length of (standard) battles with the the $H'$ family of fast growing functions.

@@

\begin{Coqsrc}
Theorem battle_length_std_Hardy (alpha : E0) :
  alpha <> Zero ->
  forall k , 1 <= k -> exists l: nat,  
       H'_ alpha k - k <= l /\
       battle_length standard k (iota (cnf alpha)) l.    
\end{Coqsrc}

@@

%%% TODO : Check the statement !
%
% \index{hydras}{Exercises}
% \begin{exercise}
% Let us quote a theorem from ~\cite{KS81} (page 297).

% \begin{quote}
% \begin{align*}
%   H'_{\omega^\alpha}(n+1) &\geq F_{\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) \\
%  F_{\alpha}(n+1) &\geq H'_{\omega^\alpha}(n) \quad (n\geq 1, \alpha<\epsilon_0) 
% \end{align*}
% Thus $H'_{\omega^\alpha}$ and $F_{\alpha}$ have essentially the same order of growth.

% \end{quote}

%  But, before trying to prove these facts, look at the definition of function $H$ in Ketonen and Solovay's paper ! Is it really the same as the definition we quote from Pr{\H o}mel's chapter~\cite{Promel2013},
% whereas \cite{KS81} define $H_\alpha(n)$ as ``the least integer $k$ such that $[n,k]$ is $\alpha$-large''. Thus, it may be useful to adapt the statement above.
